Carbonate chemistry at Ny-Ålesund

### 2023-03-13 awipev-CO2.Rmd
further figs refinement

pco2_delta_plt <- tmp_delta %>%
  ggplot(aes(x = montht, y = pco2_remos_ctd_temp_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(limits = c(-46,10), expand = c(0, 0)) +
  labs(y = latex2exp::TeX("$pCO_2$ at in situ temperature ($\\mu atm$)", bold = TRUE)) +
  geom_col(fill = "blue") +
  geom_hline(yintercept = 0) +
  annotate(
    "text",
    x = 10.5,
    y = 6,
    parse = TRUE,
    label = "pCO['2 surface']>pCO['2 deep']", size = size_labs/.pt, fontface = "bold"
  ) +
  annotate(
    "text",
    x = 10.5,
    y = -40,
    parse = TRUE,
    label = "pCO['2 surface']<pCO['2 deep']", size = size_labs/.pt, fontface = "bold"
  ) +
  Mytheme(size_labs = size_labs) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.title.x = element_text(face = "bold", size = size_labs),
        axis.title.y = element_text(face = "bold", size = size_labs)
  )
pco2_delta_plt
# we look at pCO2 normalized at 4°C at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                #pHint_tot_sf,
                pco2_remos_ctd_temp,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(pco2_remos_4 = seacarb::pCO2insi(pCO2lab = pco2_remos_ctd_temp, Tlab = temp_insitu_ctd, SST = 4)
  )
# add pco2_remos_4 to d_all
d_all <-
  dplyr::left_join(d_all, dplyr::select(tmp, datetime, pco2_remos_4), by = "datetime") %>%
  dplyr::select(sort(everything()))

@@

# monthly 
taka_month <- taka %>%
  group_by(montht) %>%
  dplyr::mutate(
    pco2_thermal =
      mean(pco2_fb_insi_temp, na.rm = TRUE) * exp(0.0423 * (
        temp_insitu_11m - mean(temp_insitu_11m, na.rm = TRUE)
      )),
    pco2_nonthermal = pco2_fb_insi_temp * exp(0.0423 * (
      mean(temp_insitu_11m, na.rm = TRUE) - temp_insitu_11m
    ))
  ) %>%
  dplyr::summarise(
    thermal = max(pco2_thermal, na.rm = TRUE) - min(pco2_thermal, na.rm = TRUE),
    nonthermal = max(pco2_nonthermal, na.rm = TRUE) - min(pco2_nonthermal, na.rm = TRUE),
    thermal_nonthermal = thermal / nonthermal
  ) %>%
  ungroup()
# monthly plot
taka_month_plt <- taka_month %>% 
  ggplot() +
  geom_point(aes(x = montht, y = thermal_nonthermal), col="blue", size = 2) +
  labs(x="Month", y= latex2exp::TeX("Thermal/non-thermal control of $pCO_{2}",
                                    bold = TRUE))  +
  geom_hline(yintercept = 1, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 14) + 
  theme(aspect.ratio = 1,
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold")) +
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
taka_month_plt
ggsave(filename = "figures/taka_month.png", plot = taka_month_plt, width = 25, height = 25, units = "cm")
ggsave(filename = "figures/essd/taka_month.pdf", plot = taka_month_plt, width = 25, height = 25, units = "cm")

@@

# For paper
size_point <- 1.5 # size of data points
alpha <- 1 #transparency
size_labs <- 14 # font size labels
# temp_insitu_11m
ts_air_sea <- air_sea_month %>%
  ggplot(aes(x = montht, y = f_air_sea / 1000), na.rm = TRUE) +
  geom_col(colour="blue", fill = "blue",  alpha = alpha, stat="identity") +
  lims(y = c(-3, 0)) +
  geom_hline(yintercept = 0) +
  labs(title=NULL, 
       x="Month", 
       y=latex2exp::TeX("Air-sea CO$_2$ flux (mol $m^{-2}$ month$^{-1}$)", bold = TRUE)) +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs))
ts_air_sea
ggsave(filename = "figures/ts_air_sea.png", plot = ts_air_sea, width = 20, units = "cm")
ggsave(filename = "figures/essd/ts_air_sea.pdf", plot = ts_air_sea, width = 20, units = "cm")
#rm(p, p_f_air_sea, p_f_air_sea_d, p_f_air_sea_m)
```

>>>>

pco2_delta_plt <- tmp_delta %>%
  ggplot(aes(x = montht, y = pco2_remos_ctd_temp_med_delta)) +
  scale_x_discrete("Month", expand = c(0, 0)) +
  scale_y_continuous(limits = c(-46,10), expand = c(0, 0)) +
  labs(y = latex2exp::TeX("$pCO_2$ at in situ temperature ($\\mu atm$)", bold = TRUE)) +
  geom_col(fill = "blue") +
  geom_hline(yintercept = 0) +
  annotate(
    "text",
    x = 10.5,
    y = 6,
    parse = TRUE,
    label = "pCO['2 surface']>pCO['2 deep']", size = size_labs/.pt, fontface = "bold"
  ) +
  annotate(
    "text",
    x = 10.5,
    y = -40,
    parse = TRUE,
    label = "pCO['2 surface']<pCO['2 deep']", size = size_labs/.pt, fontface = "bold"
  ) +
  Mytheme(size_labs = size_labs) +
  theme(
    plot.margin = margin(t = 0, r = 0.2, b = 0, l = 0, unit = "cm"),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.title.x = element_text(face = "bold", size = size_labs),
    axis.title.y = element_text(face = "bold", size = size_labs)
  )
pco2_delta_plt
# we look at pCO2 normalized at 4°C at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                #pHint_tot_sf,
                pco2_remos_ctd_temp,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(pco2_remos_4 = seacarb::pCO2insi(pCO2lab = pco2_remos_ctd_temp, Tlab = temp_insitu_ctd, SST = 4)
  )
# add pco2_remos_4 to d_all
d_all <-
  dplyr::left_join(d_all, dplyr::select(tmp, datetime, pco2_remos_4), by = "datetime") %>%
  dplyr::select(sort(everything()))

@@

# monthly 
taka_month <- taka %>%
  group_by(montht) %>%
  dplyr::mutate(
    pco2_thermal =
      mean(pco2_fb_insi_temp, na.rm = TRUE) * exp(0.0423 * (
        temp_insitu_11m - mean(temp_insitu_11m, na.rm = TRUE)
      )),
    pco2_nonthermal = pco2_fb_insi_temp * exp(0.0423 * (
      mean(temp_insitu_11m, na.rm = TRUE) - temp_insitu_11m
    ))
  ) %>%
  dplyr::summarise(
    thermal = max(pco2_thermal, na.rm = TRUE) - min(pco2_thermal, na.rm = TRUE),
    nonthermal = max(pco2_nonthermal, na.rm = TRUE) - min(pco2_nonthermal, na.rm = TRUE),
    thermal_nonthermal = thermal / nonthermal
  ) %>%
  ungroup()
# monthly plot
taka_month_plt <- taka_month %>%
  ggplot() +
  geom_point(aes(x = montht, y = thermal_nonthermal),
             col = "blue",
             size = 2) +
  labs(
    x = "Month",
    y = latex2exp::TeX("Thermal/non-thermal control of $pCO_{2}",
                       bold = TRUE)
  )  +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() + Mytheme(size_labs = 14) +
  scale_x_discrete(breaks = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"), 
                   labels = c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", ""))+
  coord_fixed(ratio = 1,
              expand = TRUE,
              clip = "on") +
  theme(
    plot.margin = margin(
      t = 0.2,
      r = 0.2,
      b = 0.2,
      l = 0.2,
      unit = "cm"
    ),
    aspect.ratio = 1,
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold")
  )
taka_month_plt
ggsave(filename = "figures/taka_month.png", plot = taka_month_plt, width = 25, height = 25, units = "cm")
ggsave(filename = "figures/essd/taka_month.pdf", plot = taka_month_plt, width = 12, height = 12, units = "cm")

@@

# For paper
size_point <- 1.5 # size of data points
alpha <- 1 #transparency
size_labs <- 14 # font size labels
# temp_insitu_11m
ts_air_sea <- air_sea_month %>%
  ggplot(aes(x = montht, y = f_air_sea / 1000), na.rm = TRUE) +
  geom_col(colour="blue", fill = "blue",  alpha = alpha) +
  lims(y = c(-3, 0)) +
  geom_hline(yintercept = 0) +
  labs(title=NULL, 
       x="Month", 
       y=latex2exp::TeX("Air-sea CO$_2$ flux (mol $m^{-2}$ month$^{-1}$)", bold = TRUE)) +
  Mytheme(size_labs = size_labs) +
  theme(plot.margin = margin(t = 0.2, r = 0.2, b = 0.2, l = 0.2, unit = "cm"),
        axis.title.x = element_text(face = "bold", size = 14),
        axis.title.y = element_text(face = "bold", size = 14),
        axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=14)
        ) +
  scale_x_discrete(breaks = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"), 
                   labels = c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", ""))
ts_air_sea
ggsave(filename = "figures/ts_air_sea.png", plot = ts_air_sea, width = 20, units = "cm")
ggsave(filename = "figures/essd/ts_air_sea.pdf", plot = ts_air_sea, width = 20, units = "cm")
#rm(p, p_f_air_sea, p_f_air_sea_d, p_f_air_sea_m)
```
